---
title: "HPAM 7660 Data Assignment 1"
author: "Rebecca Letsinger"
date: "February 6, 2024"
output: pdf_document
---

```{r}
install.packages("dplyr")
install.packages("readr")
install.packages("tidry")
install.packages("nycflights13")
install.packages("fivethirtyeight")
```

```{r}
library(dplyr)
library(readr)
library(tidyr)
library(nycflights13)
library(fivethirtyeight)
```
Question 2:
```{r}
glimpse(drinks)
```

Question 3:
```{r}
?drinks
```

Question 4: In the context of R and data analysis, a data set is considered "tidy" if it follows a specific structure that makes it easy to perform data analysis and visualization. This includes column, rows, and observational units within those columns and rows. Wide data has each row representing a unique observation, but multiple variables are spread across columns. Long data has each variable and its corresponding values are stored in separate rows. 

Question 5: This data is not tidy because it is not broken down into servings per spirit. The values spread throughout various columns. Making it tidy will show a more complete breakdown of the table and avoid over counting.

```{r}
drinks_smaller <- drinks %>% 
  filter(country %in% c("USA", "China", "Italy", "Saudi Arabia")) %>% 
  select(-total_litres_of_pure_alcohol) %>% 
  rename(beer = beer_servings, spirit = spirit_servings, wine = wine_servings)
drinks_smaller
```

Question 6:
```{r}
drinks_smaller %>% 
  pivot_longer(names_to = "type", 
               values_to = "servings", 
               cols = beer:wine)
```
               
Question 7:
```{r}
View("drinks_smaller")
```

Question 8:
```{r}
glimpse(airline_safety)
help("airline_safety")
library(dplyr)
library(tidyr)
library(stringr)
airline_safety_tidy <- airline_safety %>%
  pivot_longer(-c(airline, incl_reg_subsidiaries, avail_seat_km_per_week), 
    names_to = "type", values_to = "count") %>%
  mutate(
    period = str_sub(type, start=-5),
    period = str_replace_all(period, "_", "-"),
    type = str_sub(type, end=-7)
  )
```

```{r}
airline_safety_smaller_tidy <-airline_safety_smaller %>%
  pivot_longer(cols = c(fatalities_85_99,                         fatalities_00_14),
              names_to = "fatalities",
              values_to = "count" )
```

```{r}
View("airline_safety_smaller_tidy")
```

Question 9:
```{r}
dem_score <- read_csv("https://moderndive.com/data/dem_score.csv")
dem_score
```

Question 10:
```{r}
View("dem_score")  
```

Question 11:
This is not tidy. If this were to be in a tidy format, the years would be separated in one column called which would describes all the years as one variable.